---
title: "ESM 206 Lab 6"
author: "Alia Ajina"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank Based Test Example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test).

H0: Equal ranks (or medians)
HA: Ranks are nor equal (or, medians are not equal)

### Create mock samples

`set.seed()` to create 'pseudorandom, but reproducible samples'

```{r}
set.seed(1414) #if we use set.seed it will set the starting number and if it is all the same every R code will get the same number sequence
gp_1 <- sample.int(20, size = 15, replace = TRUE) # numbers from 1 to 20, 15 different numbers

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE) # Replace makes it a random sample
```

Take a quick look at the distributions using `hist()`

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the ardument 'paired = "TRUE"' within the t.test() function.
```

If they come from populations with the same mean, then there's a 19.8 percent chance of finding random sample means that are at least as different as the ones I found from my samples.

This p-value means that IF our samples are drawn from populations with the same mean (ie., if the null hypothesis is true), there is a `r round(ttest$p.value,2)` probability of finding sample means *at least as different* as mine by random chance. 
(command enter allows you to run the value you're calculating IE .2)

[A more comprehensive explanation of difference between groups... actually means, data spread, effect size]... there is no significant difference (t(df) t-statistic value, p =)

You should say "there is no significant difference" never say "we fail to reject the null hypothesis) she doesn't like this language

BUT make sure you don't say accept-- say we fail to reject or we retain the null hypothesis, she thinks this is a stupid conversation

### Mann Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that IF our samples are drawn from populations with the same median (ie., if the null hypothesis is true), then there is a 28% chance of finding samples medians *at least as different* as ours by random chance. 

No significant difference in medians (Mann-Whitney) U; U = `r mwu$statistic`, p = `r round(mwu$p.value, 2)`)

Example tect for reporting: "The medians of group A (M = 22.4 g) and group B (M = 31.2 g) did not differ significantly (Mann Whitney U, ....)."

### Simple linear regression

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense.

Use the `penguins` data from {palmerpenguins}, exploring flipper length (x) versus body mass(y).

```{r}
# Use View(penguins) to look at it
# Make an explroatory scatterplot of penguin flipper length (x-axis) versus body mass( y-axis)

ggplot(data = penguins, 
       aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point() +
  geom_smooth(method = "lm")

# If you present something with linear modeling you are saying that you understand the math behind the creation of this trend line, you better understand what that R^2 value means, you better be able to explain it

```

### Find the Model Estimates (coefficients)

Use the `lm()` function for linear model.

when I do body_mass_g ~ <<<< that is a tilde, not a minus sign
it says run body mass g as a function of flipper, body mass g is the dependent variable

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

 So this says, it's y=mx+b, body_mass_g (y) = 49.69(g/mm)(flipper_length_mm) - 5780.8 (g)
 49.69/1 (g/mm), the slope means on average for a 1 mm increase in flipper length, there is a 49.69 g increase in body mass on average
 
the y-intercept is the mass of penguin with no flippers, a flipper length of 0, based on our model that is the expected average mass of penguins that have a flipper length of 0mm
 - this is important in giving us our correct starting point
 - a negative weight doesnt exist, it's not logical in real life
 - the y-intecept on it's own is sometimes not meaningful to interpret
 
For each coeeficient value it does a statistical hypothesis where the value of those coefficients is equal to 0
 - we have significantly non-0 coefficients 
 
 Model coefficients: `r penguin_lm$coefficients`
 Flipper length coefficient :`r penguin_lm$coefficients[2]`
 
 [] square brackets are called indexing, so ^ that calls the second coefficient only, it signifies what you want to call and show up in your knitted doc
 
### Use `broom::tidy()` to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy # this gets us a lot of the way towards a finalized table of regression outcomes but it's not finalized yet, there are other ways to change it
```

### Use `broom::glance()` to get output of model fit & statistical significance

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```

### Use equatiomatic::extract_eq() to return model in LaTeX (a language used to write mathematical equations)
the :: means in this package (equatiomatic) use extract_eq() 

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

### Use plot() function to automatically produce diagnostic pools

```{r}
plot(penguin_lm)
```

